Trauma treatment delay

### 2023-11-30 manuscriptlongversion.Rmd
text

The outcome is delayed treatment as identified by the multidisciplinary
review board and recorded in the trauma care quality database. The
independent variables are gender, highest hospital care level, GCS on
arrival, respiratory rate on arrival, systolic blood pressure on
arrival, if the patient was intubated on arrival, whether they arrived
during weekend or working hours, ISS and age. Of these variables, GCS,
respiratory rate, systolic blood pressure, intubation, ISS and age are
used as continuous variables in the analysis

@@

The study size is all of the eligible patients that was treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs. During this period, 72
patients were deemed to have received delayed treatment.

@@

Rstudio was used to perform the logistical regression. The code was
written to clean and prepare the data from the database. The data was
processed to have the variables customized to fit into the logistical
regression model, such as making them into factors, renaming outcomes in
the database and removing unknown values. The variables were either
binary or continuous.

Using the one in ten rule, 7 independent variables were used for the
logistical regression given that we have 72 patients with delayed
treatment. The one in ten rule states that for every variable examined,
10 events are needed in the group. In this case, this means that because
we have 72 patients with the examined outcome, we can have 7 variables.
As we had more variables we wanted to use in the regression model, we
decided to run two logistical regression models. One with gender, age,
ISS, emergency procedure, SBP, RR and GCS. The other one with gender,
age, ISS, highest care level, intubation, whether they arrived during
work hours and whether they arrived on a weekday.

@@

The logistical regression is run through Rstudio using the glm command.
To make the tables, the package gtsummary was used to visualize the
adjusted and unadjusted regression, then merging the two tables into
one.

@@

As were running two logistical regression models with different
variables and a missing value in a variable led to exclusion, the two
groups did not have the same amount of participants. In the first
logistical regression, 1,261 patients had missing data in one or more of
the analysed parameters which left 7,051 individuals group, 69 of which
received delayed treatment. In the second logistical regression, 1,038
patients had missing data in one or more of the analysed parameters
which left 7,274 individuals group, 70 of which received delayed
treatment.

@@

The analysis shows that when adjusted for, most variables do not have a
correlation with receiving delayed treatment. Three variables, ISS, RR
and GCS, had a correlation with receiving delayed treatment when not
adjusted for. However, when adjusted for, the correlation dissapeared.
In the case of emergency procedure, there was a correlation for all
emergency procedures performed and having a lowered incidence of delayed
treatment among the patients, compared with not receiving any emergency
measures performed

@@

## Discussion

>>>>

The outcome is delayed treatment as identified by the multidisciplinary
review board and recorded in the trauma care quality database. The
independent variables are gender, GCS on arrival, respiratory rate on
arrival, systolic blood pressure on arrival, number of injuries, ISS and
age. Of these variables, GCS, respiratory rate, systolic blood pressure,
number of injuries, ISS and age are used as continuous variables in the
analysis. Gender is used as a categorical variable.

@@

The study size is all of the eligible patients that was treated at the
hospital between 2012 and 2022. Starting 2013 only a subgroup of
patients were screened for OFIs, but starting 2017 all patients that are
included in the database are screened for OFIs. Out of the 14,022
patients in the database, 7,378 were included in the study, with the
rest being excluded due to either not being screened for ofi or missing
values in the examined variables. 72 patients were deemed to have
received delayed treatment. However, only 70 patients with the
registered outcome were included in the study due to missing values
among two of the patients.

@@

Rstudio was used to perform the logistical regression. The code was
written to clean and prepare the data from the database. The data was
processed to have the variables customized to fit into the logistical
regression model, such as making them into factors, renaming outcomes in
the database and removing unknown values. The variables were either
binary or continuous. This code was used on scrambled data at first so
to avoid selection bias as much as possible. When the code was finished,
it was run on the unscrambled data.

Using the one in ten rule, 7 independent variables were used for the
logistical regression given that we have 72 patients with delayed
treatment. The one in ten rule states that for every variable examined,
10 events are needed in the group. In this case, this means that because
we have 72 patients with the examined outcome, we can have 7 variables.

@@

The logistical regression is run through Rstudio using the glm command.
To make the tables, the package gtsummary was used to visualize the
adjusted and unadjusted regression.

@@

As we were running the logistical regression, further were excluded from
the study due to missing data. Among the 8,312 patients screened for
ofi, 7,378 were included in the study. 934 patients were excluded from
the study due to missing values in one or more variables.

@@

The analysis shows that some variables seem to have a correlation with
patients receiving delayed treatment. ISS and respiratory rate both have
a trend that patient with higher values seem to have a lower incidence
of delayed treatment, both unadjusted and adjusted. GCS had the opposite
correlation, with lower GCS often having lower incidence of delayed
treatment, both unadjusted and adjusted. Number of injuries had the
correlation of higher number of injuries having lower incidence of
delayed treatment unadjusted, but that correlation was not seen when
adjusted for.

@@

# Discussion
